---
title: "README"
output:
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# Load packages
library(tidyverse)
# set pander table-layout options
library(pander)
panderOptions('table.alignment.default', function(df)
  ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```




## Code for the main analysis of the paper "Waning Immunity After the BNT162b2 Vaccine in Israel"

### Data

The data for the main analysis of documented infection can be found in the file `pos_data_days11-31_7.csv` and for severe illness in `sev_data_days11-31_7.csv`. 
To read the data into the R Statistical Software use the following code:
```{r, echo = T}

dat_positive <- read_csv("pos_data_days11-31_7.csv") %>%
  mutate(`Vacc Period` = factor(`Vacc Period`,levels = c("JanB", "FebA", "FebB", "MarA", "MarB", "Apr", "May")), 
         `Epi Week` = factor(`Epi Week`),
         Sector = factor(Sector, levels = c("General Jewish", "Arab","ultra-Orthodox Jewish")),
        `Past PCR tests` =  factor(`Past PCR tests`, levels = c("0","1","2+")),
         N_person = as.numeric(N_person),
        Positive =  Rate_Positive_1K*N_person/1000)

dat_severe <- read_csv("sev_data_days11-31_7.csv") %>%
  mutate(`Vacc Period` = factor(`Vacc Period`,levels = c("Jan", "Feb", "Mar", "Apr-May")), 
         `Epi Week` = factor(`Epi Week`),
         Sector = factor(Sector, levels = c("General Jewish", "Arab","ultra-Orthodox Jewish")),
        `Past PCR tests` =  factor(`Past PCR tests`, levels = c("0","1","2+")),
         N_person = as.numeric(N_person),
       Severe =  Rate_Severe_1K*N_person/1000)


```

### Analysis

To run the main analysis of the paper for documented infection use the following code. 

```{r analysis, echo = T}
formula_positive <- as.formula("Positive ~  Age +
                         Gender +
                         `Epi Week` +
                        `Past PCR tests` +
                         Sector +
                          `Vacc Period`:Age +
                         offset(log(N_person))")


formula_severe <- as.formula("Severe ~   Age +
                         Gender +
                         `Epi Week` +
                        `Past PCR tests` +
                         Sector +
                          `Vacc Period`:Age +
                         offset(log(N_person))")

analysis_positive <- glm(formula_positive, family="poisson", data=dat_positive)
analysis_severe <- glm(formula_severe, family="poisson", data=dat_severe)

```

### Confidence Intervals

To calculate 95% confidence interval for the rate ratio of January and May for confirmed infections in the  60 and above age group use the following code:
```{r CIpos, echo = T}


df <- broom::tidy(analysis_positive) %>%
  filter(term == "Age60+:`Vacc Period`May") %>%
  mutate(est =  round(exp(-estimate),1),
         lower = round(exp(-estimate-1.96*std.error),1),
         upper = round(exp(-estimate+1.96*std.error),1),
         `Confidence Interval: Positive`= paste0(est," 95% CI: [",lower,", ",upper,"]")) %>%
  select(`Confidence Interval: Positive`)

pander::pander(df)



```

To calculate 95% confidence interval for the rate ratio of January and March for severe illness in the  60 and above age group use the following code:
```{r CIsev, echo = T}


df <- broom::tidy(analysis_severe) %>%
  filter(term == "Age60+:`Vacc Period`Mar") %>%
  mutate(est =  round(exp(-estimate),1),
         lower = round(exp(-estimate-1.96*std.error),1),
         upper = round(exp(-estimate+1.96*std.error),1),
         `Confidence Interval: Severe`= paste0(est," 95% CI: [",lower,", ",upper,"]")) %>%
  select(`Confidence Interval: Severe`)

pander::pander(df)
```


